---
title: "Coaching and organisational strucutures in eSports"
output:
  pdf_document: default
  html_document: default
---

***

## Load data

```{r load data}
## Directory and file settings
dir <- "data"
filename <- "160913_Coaching-and-organisational-structures-in-eSports.csv"

## Read data
dat <- read.csv2(file.path(dir, filename), 
                 header = T, 
                 sep = ",",
                 stringsAsFactors = F)
```

Data structure:

```{r data structure}
str(dat)
```

***

## Recode variable names

```{r varname}
## Save old variable names
colnames_old <- colnames(dat)

## Prepare new variable names
colnames_new <- c(
  "time",
  "age",
  "time_play",
  "time_spect",
  "role",
  "team_solo",
  "rec_coach",
  "games_playcompete",
  "games_important",
  "region",
  "region_reason",
  "importance_overall",
  "importance_manager",
  "importance_manager_txt",
  "importance_coach",
  "importance_coach_txt",
  "importance_analyst",
  "importance_analyst_txt",
  "future"
)

## Recode variable names
colnames(dat) <- colnames_new

## For testing purpose
#print(data.frame("Old varnames" = colnames_old, "New varnames" = colnames_new))
```

***

## Data cleaning

__Values ranges__

* Age (optional):
    + "under 12"
    + "12 - 15"
    + "16 - 18"
    + "19 - 21"
    + "21 - 24" __WHY 21 AGAIN????__
    + "25 - 27"
    + "28 - 30"
    + "31 - 35"
* Role:
    + Player
    + Manager
    + Analyst
    + Coach
    + Scout
    + Caster
    + Sonstiges: 
* Preference:
    + "Solo"
    + "Team"
* Importance: 1-10 (not important - essential)
* Future: 1- 10 (not important - essential)
* Time (play, watch):
    + "under 1 hour"
    + 1-3 hours"
    + "3-5 hours"
    + "5-7 hours"
    + "8-10 hours"
* Received coaching:
    + "currently getting coached"
    + "got coached in the last 3 month"
    + "got coached in the last 6 month"
    + "got coached in the last year"
    + "got coached but way earlier"
    + "didn't receive any coaching"
* Games (play ~ multiple):
    + "None"
    + "League of Legends"
    + "Counterstrike GO"
    + "Dota 2"
    + "Smite"
    + "Hearthstone"
    + "World of Warcraft"
    + "Overwatch"
    + "Sonstiges"
* Region:
    + "Europe"
    + "North America"
    + "South America"
    + "Korea"
    + "China"
    + "Asia (without China and Korea)"
    + "Australia"
    + "Afrika"
    + "Sonstiges"
* Reason(multiple):
    + "mentality of the players"
    + "time investment"
    + "structural organisation"
    + "coaching"
    + "popularity of eSports in the region"
    + "salaries and investments"
    + "Sonstiges"

***

### Remove trolls

Identified trolls (to be removed:
* Timestamp: 29.08.2016 20:19:10

```{r remove troll}
## Remove troll
case2rm <- which(dat$time == "29.08.2016 20:19:10")
dat <- dat[-case2rm, ]
```

***

### Question: What's your role in eSports?

Given answers in free text field "Sonstiges"

Recode scheme:

* "Teamorganizer" to "Manager""
* "Viewer" to "Player" or remove case
    + __Final decision: Remove case__
* "Casual Player" to "Player""
* "Many of the above"
    + __STILL UNCLEAR__

```{r role processing}
## Identify misc cases
casefinder <- 
  !dat$role %in% c("Player", "Manager", "Analyst", "Coach", "Scout", "Caster")
print(dat$role[casefinder])

## Recode into new variable role2
dat$role2 <- dat$role
dat$role2[which(dat$role2 == "Teamorganizer")] <- "Manager"
dat$role2[which(dat$role2 == "Casual Player")] <- "Player"

## Remove cases
case2rm <- which(dat$role == "Viewer")
dat <- dat[-case2rm, ]

case2rm <- which(dat$role == "Many of the above, including Org Owner")
dat <- dat[-case2rm, ]

case2rm <- which(grepl("None", dat$role))
dat <- dat[-case2rm, ]
case2rm <- which(dat$role == "/")
dat <- dat[-case2rm, ]
```

***

### Questions regarding games (play/spectate):

Recode into new variables-per-game

```{r}
library(dplyr)

## Games played competetively
dat <- 
  dat %>%
  ## Workaround: Re-initialise as data.frame
  data.frame() %>% 
  mutate(games_playcompete_none = ifelse(grepl("None", games_playcompete), 1, 0),
         games_playcompete_lol = ifelse(grepl("League of Legends", games_playcompete), 1, 0),
         games_playcompete_csgo = ifelse(grepl("Counterstrike GO", games_playcompete), 1, 0),
         games_playcompete_dota2 = ifelse(grepl("Dota 2", games_playcompete), 1, 0),
         games_playcompete_smite = ifelse(grepl("Smite", games_playcompete), 1, 0),
         games_playcompete_heartstone = ifelse(grepl("Hearthstone", games_playcompete), 1, 0),
         games_playcompete_wow = ifelse(grepl("World of Warcraft", games_playcompete), 1, 0),
         games_playcompete_overwatch = ifelse(grepl("Overwatch", games_playcompete), 1, 0),
         games_playcompete_skill = ifelse(grepl("S.K.I.L.L.", games_playcompete), 1, 0),
         games_playcompete_startcraft2 = ifelse(grepl("Starcraft 2", games_playcompete), 1, 0),
         games_playcompete_bf34 = ifelse(grepl("Battlefield 3/4", games_playcompete), 1, 0),
         games_playcompete_codbo3 = ifelse(grepl("Call of Duty: Black Ops 3", games_playcompete), 1, 0),
         games_playcompete_codmw2 = ifelse(grepl("call of duty modern warfare 2", games_playcompete), 1, 0),
         games_playcompete_rss = ifelse(grepl("Rainbow Six Siege", games_playcompete), 1, 0)
         )

## Region reason
dat <- 
  dat %>%
  ## Workaround: Re-initialise as data.frame
  data.frame() %>% 
  mutate(region_reason_mentality = ifelse(grepl("mentality of the players", region_reason), 1, 0),
         region_reason_time = ifelse(grepl("time investment", region_reason), 1, 0),
         region_reason_structuralorg = ifelse(grepl("structural organisation", region_reason), 1, 0),
         region_reason_coaching = ifelse(grepl("coaching", region_reason), 1, 0),
         region_reason_popularity = ifelse(grepl("popularity of eSports in the region", region_reason), 1, 0),
         region_reason_salaries = ifelse(grepl("salaries and investments", region_reason), 1, 0)
         )

## Age
dat$age[dat$age == ""] <- NA

## Coaching
dat$rec_coached10 <- as.character(dat$rec_coach)
dat$rec_coached10[which(grepl("didn't", dat$rec_coached10))] <- 0
dat$rec_coached10[which(dat$rec_coached10 != "no coaching")] <- 1

dat$rec_coach_currently10 <- as.character(dat$rec_coach)
dat$rec_coach_currently10[which(grepl("didn't", dat$rec_coach_currently10))] <- 0
dat$rec_coach_currently10[which(grepl("currently", dat$rec_coach_currently10))] <- 1

dat$rec_coach2 <- as.character(dat$rec_coach)
dat$rec_coach2[which(grepl("didn't", dat$rec_coach2))] <- "never been coached"
dat$rec_coach2[which(grepl("currently", dat$rec_coach2))] <- "currently coached"
dat$rec_coach2[which(grepl("got", dat$rec_coach2))] <- "been coached"
```


***

## Analysis

### Research questions of interest ###

* importance_coach vs. rec_coach

* importance_X vs. time_play
* importance_X vs. time_spect
* importance_X vs. role
* importance_X vs. team_solo

* future vs. time_play
* future vs. time_spect
* future vs. role
* future vs. team_solo

***

### How much time do you play games per day?

```{r plot time vs. importance + age}

dat$time_play <- factor(dat$time_play)
print(levels(dat$time_play))
dat$time_play <- factor(dat$time_play, levels = levels(dat$time_play)[c(6, 1:5)])

dat$time_spect <- factor(dat$time_spect)
print(levels(dat$time_spect))
dat$time_spect <- factor(dat$time_spect, levels = levels(dat$time_spect)[c(6, 1:5)])

library(ggplot2)

var4y_list <- c(
  "importance_overall",
  "importance_manager",
  "importance_coach",
  "importance_analyst"
)

for(var4y in var4y_list) {
  
  set.seed(42)
  plotdat <- 
    ggplot() + 
    geom_boxplot(data = dat,
                 aes_string(x = "time_play",
                            y = var4y),
                 alpha = 0.5) + 
    geom_point(data = dat,
               aes_string(x = "time_play",
                          y = var4y,
                          colour = "age"),
               position = position_jitter(w = 0.25, h = 0.25),
               size = 2.5) + 
    scale_y_continuous(limits = c(1, 10), breaks = seq(1, 10, 1)) + 
    #theme(axis.text.x = element_text(angle = 45)) +
    ggtitle(paste("time play vs.", var4y))
  plot(plotdat)
  
  set.seed(42)
  plotdat <- 
    ggplot() + 
    geom_boxplot(data = dat,
                 aes_string(x = "time_spect",
                            y = var4y),
                 alpha = 0.5) + 
    geom_point(data = dat,
             aes_string(x = "time_spect",
                        y = var4y,
                        colour = "age"),
             position = position_jitter(w = 0.25, h = 0.25),
               size = 2.5) + 
    scale_y_continuous(limits = c(1, 10), breaks = seq(1, 10, 1)) +
    #theme(axis.text.x = element_text(angle = 45)) +
    ggtitle(paste("time spect vs.", var4y))
  plot(plotdat)
}
  
```


### Role

```{r plot role vs. importance}

for(var4y in var4y_list) {
  set.seed(42)
  plotdat <- 
    ggplot() +
    geom_boxplot(data = dat,
                 aes_string(x = "role2",
                            y = var4y),
                 alpha = 0.5) +
    geom_point(data = dat,
               aes_string(x = "role2",
                          y = var4y,
                          colour = "age"),
               position = position_jitter(w = 0.25, h = 0.25),
               size = 2.5) + 
    scale_y_continuous(limits = c(1, 10), breaks = seq(1, 10, 1)) + 
    theme(axis.text.x = element_text(angle = 45)) +
    ggtitle(paste("role vs.", var4y))
  plot(plotdat)
}
```

```{r plot role vs. future}
set.seed(42)
  plotdat <- 
    ggplot() + 
    geom_boxplot(data = dat,
                 aes(x = role2,
                     y = future),
                 alpha = 0.5) +
    geom_point(data = dat,
               aes(x = role2,
                   y = future,
                   colour = age),
               position = position_jitter(w = 0.25, h = 0.25),
               size = 2.5) + 
    scale_y_continuous(limits = c(1, 10), breaks = seq(1, 10, 1)) + 
    theme(axis.text.x = element_text(angle = 45)) +
    ggtitle(paste("role vs. future"))
  plot(plotdat)
```

```{r coach vs. future}
set.seed(42)
plotdat <- 
  ggplot() + 
  geom_boxplot(data = dat,
               aes(x = rec_coach,
                   y = future),
               alpha = 0.5) +
  geom_point(data = dat,
             aes(x = rec_coach,
                 y = future,
                 colour = age),
             position = position_jitter(w = 0.25, h = 0.25),
             size = 2.5) + 
  scale_y_continuous(limits = c(1, 10), breaks = seq(1, 10, 1)) + 
  theme(axis.text.x = element_text(angle = 45)) +
  ggtitle(paste("Received coaching vs. future"))
plot(plotdat)
  
set.seed(42)
plotdat <- 
  ggplot() + 
  geom_boxplot(data = dat,
               aes(x = rec_coach2,
                   y = future),
               alpha = 0.5) +
  geom_point(data = dat,
             aes(x = rec_coach2,
                 y = future,
                 colour = age),
             position = position_jitter(w = 0.25, h = 0.25),
             size = 2.5) + 
  scale_y_continuous(limits = c(1, 10), breaks = seq(1, 10, 1)) + 
  theme(axis.text.x = element_text(angle = 45)) +
  ggtitle(paste("Received coaching vs. future"))
plot(plotdat)
```


```{r coaching}
dat$rec_coach <- factor(dat$rec_coach)
dat$rec_coach <- factor(dat$rec_coach, levels = levels(dat$rec_coach)[c(2, 3, 6, 5, 4, 1)])

for(var4y in var4y_list) {
  set.seed(42)
  plotdat <- 
    ggplot() +
    geom_boxplot(data = dat,
                 aes_string(x = "rec_coach",
                            y = var4y),
                 alpha = 0.5) +
    geom_point(data = dat,
               aes_string(x = "rec_coach",
                          y = var4y,
                          colour = "role2"),
               position = position_jitter(w = 0.25, h = 0.25),
               size = 2) +
    scale_y_continuous(limits = c(1, 10), breaks = seq(1, 10, 1)) +
    theme(axis.text.x = element_text(angle = 45)) +
    ggtitle(paste("Received coaching vs.", var4y))
  plot(plotdat)
}

for(var4y in var4y_list) {
  set.seed(42)
  plotdat <- 
    ggplot() +
    geom_boxplot(data = dat,
                 aes_string(x = "rec_coach2",
                            y = var4y),
                 alpha = 0.5) +
    geom_point(data = dat,
               aes_string(x = "rec_coach2",
                          y = var4y,
                          colour = "role2"),
               position = position_jitter(w = 0.25, h = 0.25),
               size = 2) +
    scale_y_continuous(limits = c(1, 10), breaks = seq(1, 10, 1)) +
    theme(axis.text.x = element_text(angle = 45)) +
    ggtitle(paste("Received coaching vs.", var4y))
  plot(plotdat)
}
```
